---
title: "R Coding Exercise"
output:
  html_document:
    toc: FALSE
---

## **LOADING AND CHECKING DATA**

### Install packages required for exercise
```{r}
#Commented out after initial install
#install.packages(c("dslabs","ggplot2","tidyverse"))
```

### Load packages
```{r}
#| warning: false
library("dslabs")
library("ggplot2")
library("tidyverse")
```

### Look at help file for gapminder data
```{r}
help(gapminder)
```

### Get an overview of data structure for gapminder
```{r}
str(gapminder)
```

### Get a summary of data in gapminder
```{r}
summary(gapminder)
```

### Determine the type of object gapminder is
```{r}
class(gapminder)
```

## **PROCESSING DATA**

### Assign African countries to a new object called africadata
```{r}
africadata<-gapminder[gapminder$continent == "Africa", ]
```

### Run str and summary on africadata to check work
#### (2907 obs and 9 vars)
```{r}
str(africadata)
summary(africadata)
```

### Create new object that contains only infant_mortality and life_expectancy
```{r}
infmort_lifeexp <- africadata[,c("infant_mortality","life_expectancy")]
```

### Run str and summary on new infmort_lifeexp object
#### (2907 obs and 2 vars)
```{r}
str(infmort_lifeexp)
summary(infmort_lifeexp)
```

### Create new object that contains only population and life_expectancy
```{r}
pop_lifeexp <- africadata[,c("population","life_expectancy")]
```

### Run str and summary on new pop_lifeexp object
#### (2907 obs and 2 vars)
```{r}
str(pop_lifeexp)
summary(pop_lifeexp)
```

## **PLOTTING**

### Plot life expectancy as a function of infant mortality 
```{r}
ggplot(data = infmort_lifeexp) + theme_bw() + 
  geom_point(mapping = aes(infant_mortality, life_expectancy), color = "red") +
  labs(title = "Life Expectancy and Infant Mortality in African Countries") +
  ylab("Life Expectancy (Years)") +
  xlab("Infant Mortality (Deaths per 1000)")
```

### Plot life expectancy as a function of population size, x-axis set to log scale
#### (Streaks of data likely clustering by year)
```{r}
ggplot(data = pop_lifeexp) + theme_bw() + 
  geom_point(mapping = aes(log10(population), life_expectancy), color = "blue") +
  labs(title = "Life Expectancy and Population Size in African Countries") +
  ylab("Life Expectancy (Years)") +
  xlab("Population Size (Deaths per 1000, Log Scale)")
```

## **MORE DATA PROCESSING**

### Figure out which years have missing data for infant mortality
#### (Years 1960 through 1981, 2016)
```{r}
table(is.na(africadata$infant_mortality)) #226 missing values
check_missing <- africadata %>% 
           select(year,infant_mortality) %>% 
           filter(is.na(infant_mortality))
table(check_missing$year)
```

### Create a new object with only the data for the year 2000 from africadata
```{r}
africadata_2000 <- africadata %>%
  filter(year == "2000")
```

### Run str and summary on new africadata_2000 object
#### (51 obs, 9 vars)
```{r}
str(africadata_2000)
summary(africadata_2000)
```

## **MORE PLOTTING**

### Plot life expectancy as a function of infant mortality (year 2000 only)
```{r}
ggplot(data = africadata_2000) + theme_bw() + 
  geom_point(mapping = aes(infant_mortality, life_expectancy), color = "red") +
  labs(title = "Life Expectancy and Infant Mortality in African Countries, Year 2000 Only") +
  ylab("Life Expectancy (Years)") +
  xlab("Infant Mortality (Deaths per 1000)")
```

### Plot life expectancy as a function of population size, x-axis set to log scale
```{r}
ggplot(data = africadata_2000) + theme_bw() + 
  geom_point(mapping = aes(log10(population), life_expectancy), color = "blue") +
  labs(title = "Life Expectancy and Population Size in African Countries, Year 2000 Only") +
  ylab("Life Expectancy (Years)") +
  xlab("Population Size (Deaths per 1000, Log Scale)")
```

## **A SIMPLE FIT**

### Using year 2000 data, fit life expectancy as the outcome and infant mortality as the predictor 
```{r}
fit1_lifeexp_infmort <- lm(life_expectancy~infant_mortality, data = africadata_2000)
```

### Using year 2000 data, fit life expectancy as the outcome and population as the predictor 
```{r}
fit2_lifeexp_pop <- lm(life_expectancy~population, data = africadata_2000)
```

### Apply summary command to both models; based on p-values what do you conclude?
```{r}
summary(fit1_lifeexp_infmort) 
summary(fit2_lifeexp_pop)
```

##### In model 1, p-value = 2.83e-08; significant relationship between life expectancy and infant mortality (as infant mortality increases, life expectancy decreases). In model 2, p-value = 0.616; no significant relationship between life expectancy and population size.


**------------------------------------------------------**

**SECTION ADDED BY IRENE CAVROS**

**------------------------------------------------------**


### Additional data plots

```{r}
#plot life expectancy as a function of GDP (year 2000 only)
ggplot(data = africadata_2000) + theme_bw() + geom_point(mapping = aes(gdp, life_expectancy), color = "red") + labs(title = "Life Expectancy and GDP in African Countries, Year 2000 Only") + ylab("Life Expectancy (Years)") + xlab("GDP")

#barplot of infant mortality in the year 2000 by country
#nicole commented out, too many countries for this plot
#ggplot(data=africadata_2000, aes (x=country, y=infant_mortality))
```

### Fit another model

```{r}
gdp_le_fit <- lm(life_expectancy~gdp, data = africadata_2000)
summary(gdp_le_fit) 

#Fit results in table
library(broom)
tidy(gdp_le_fit)
augment(gdp_le_fit)
glance(gdp_le_fit)
```

