---
title: "Module 4: Data Analysis Exercise"
author: "Nicole Luisi"
date: "`r Sys.Date()`"
output: html_document
---

This Quarto file was developed for the MADA Module 4 data analysis exercise.

**Description**: This analysis will explore mortality counts by month, year, and cause of death in the US. 

**Data Source**: CDC - [Select cause mortality data from 2014-2019](https://data.cdc.gov/NCHS/Monthly-Counts-of-Deaths-by-Select-Causes-2014-201/bxq8-mugm) has been combined with [provisional select cause mortality data from 2020-2023](https://data.cdc.gov/NCHS/Monthly-Provisional-Counts-of-Deaths-by-Select-Cau/9dzk-mvmi).

**Final Data File**: The RDS file `final_death_2015_2022.rds` is located in `\nicoleluisi-MADA-portfolio\Mod_4_Exercise\finaldata` and contains data from 2015 through 2022; rows are structured by month and year (each row contains mortality counts for one month (per year) for several causes of death. 

**Variables**: Categorical variables include *month*, *year*, and *natural_cause_prop_fct* (derived as a factor representing the proportion of all deaths due to natural causes). 
Numeric count variables are available for several causes of death by month/year: *all_cause*, *natural_cause*, *septicemia*, *malignant_neoplasm*, *diabetes_mellitus*, *alzheimer_disease*, *influenza_and_pneumonia*, *chronic_lower_respiratory*, *other_diseases_of_respiratory*, *nephritis_nephrotic_syndrome*, *symptoms_signs_and_abnormal*, *diseases_of_heart*, *cerebrovascular_diseases*, *accidents_unintentional*, *motor_vehicle_accidents*, *intentional_self_harm_suicide*, *assault_homicide*, *drug_overdose*, *covid_19_multiple_cause_of*, and *covid_19_underlying_cause*. 
Variable names are explanatory; user can continue with monthly counts or collapse to annual or other combinations. 
**Note**: *COVID-19 variables are NA for years prior to 2020.*


### **Load and Process Data**

##### **Install and load required packages**
```{r}
#| warning: false
#Commented out after initial install
#install.packages(c("here", tidyverse","gridExtra"))
library("here")
library("tidyverse")
library("gridExtra")
```

##### **Load data**
```{r}
# Monthly Counts of Deaths by Select Causes, 2014-2019
death_2014_2019 <- read.csv("https://data.cdc.gov/resource/bxq8-mugm.csv")
# Monthly Provisional Counts of Deaths by Select Causes, 2020-2023
death_2020_2023 <- read.csv("https://data.cdc.gov/resource/9dzk-mvmi.csv")
```

##### **Prepare combined file**

###### Check variables
```{r}
#| output: false
names(death_2014_2019)
head(death_2014_2019)
names(death_2020_2023)
head(death_2020_2023)
```

###### Subset variables
```{r}
# Subset variables 2014-2019
death_2014_2019_v2 <- death_2014_2019 %>% select(year, month, all_cause, natural_cause, septicemia, malignant_neoplasms, diabetes_mellitus, alzheimer_disease, influenza_and_pneumonia, chronic_lower_respiratory, other_diseases_of_respiratory, nephritis_nephrotic_syndrom, symptoms_signs_and_abnormal, diseases_of_heart, cerebrovascular_diseases, accidents_unintentional, motor_vehicle_accidents, intentional_self_harm_suicide, assault_homicide, drug_overdose)

# Rename variables in 2014-2019 file that don't match second subset  
death_2014_2019_v2 <- rename(death_2014_2019_v2, nephritis_nephrotic_syndrome = nephritis_nephrotic_syndrom)
   
# Subset variables 2020-2023                 
death_2020_2023_v2 <- death_2020_2023 %>% select(year, month, all_cause, natural_cause, septicemia, malignant_neoplasms, diabetes_mellitus, alzheimer_disease, influenza_and_pneumonia, chronic_lower_respiratory, other_diseases_of_respiratory, nephritis_nephrotic_syndrome, symptoms_signs_and_abnormal, diseases_of_heart, cerebrovascular_diseases, accidents_unintentional, motor_vehicle_accidents, intentional_self_harm_suicide, assault_homicide, drug_overdose, covid_19_multiple_cause_of, covid_19_underlying_cause)   
```

###### Concatenate all years and sort 
```{r}
# Concatenate all years of data
death_2014_2023 <- bind_rows(death_2014_2019_v2, death_2020_2023_v2)

# Sort by year, then month
death_2014_2023 <- death_2014_2023 %>%
  arrange(year, month)
```

######  Review combined file
```{r}
#| output: false
glimpse(death_2014_2023)
summary(death_2014_2023)
head(death_2014_2023)
```

###### Clean and prepare subset for analysis 
```{r}
# Limit period from 2015 through 2022
final_death_2015_2022 <- death_2014_2023 %>%
  filter(year %in% c("2015", "2016", "2017", "2018", "2019", "2020", "2021", "2022"))

# Create labeled month field
final_death_2015_2022$month_abrv <- month.abb[final_death_2015_2022$month] 

# Create new categorical var for exploratory analysis
# Calculate natural cause as proportion of all cause mortality
final_death_2015_2022$natural_cause_prop <- round(((final_death_2015_2022$natural_cause/final_death_2015_2022$all_cause)*100), digits=1)
# Make into factor
final_death_2015_2022$natural_cause_prop_fct <- factor(ifelse(is.na(final_death_2015_2022$natural_cause_prop),NA,
                                   ifelse(final_death_2015_2022$natural_cause_prop <90, 1, 
                                     ifelse(final_death_2015_2022$natural_cause_prop <= 91, 2, 3))),
                              levels=1:3,
                              labels=c("< 90%",
                                       "90%-91%",
                                       "92% +"))
# Check levels
table(final_death_2015_2022$natural_cause_prop_fct)
```

### **Explore Data**

###### Compare plots with select causes of death by month for a few years
```{r}
# Plot a subset of the data
ss1 <- subset(final_death_2015_2022, year == "2021") %>%
  select(month, drug_overdose, intentional_self_harm_suicide, assault_homicide) %>%
  gather(key = "variable", value = "value", -month)
p1 <- ggplot(ss1, aes(x = month, y = value)) + 
  geom_line(aes(color = variable), size = 1) +
  scale_color_manual(values = c("#1d95bb", "#0e9731", "#FC4E07")) +
  theme_minimal() +
  xlab("Month of Year, 2021") +
  ylab("Number of Deaths") +
  ggtitle("Number of Deaths by Select Cause, Year 2021") +
  scale_y_continuous(breaks=seq(0, 10000, 2000)) +
  scale_x_continuous(breaks=seq(1, 12, 1)) 

ss2 <- subset(final_death_2015_2022, year == "2020") %>%
  select(month, drug_overdose, intentional_self_harm_suicide, assault_homicide) %>%
  gather(key = "variable", value = "value", -month)
p2 <- ggplot(ss2, aes(x = month, y = value)) + 
  geom_line(aes(color = variable), size = 1) +
  scale_color_manual(values = c("#1d95bb", "#0e9731", "#FC4E07")) +
  theme_minimal() +
  xlab("Month of Year, 2020") +
  ylab("Number of Deaths") +
  ggtitle("Number of Deaths by Select Cause, Year 2020") +
  scale_y_continuous(breaks=seq(0, 10000, 2000)) +
  scale_x_continuous(breaks=seq(1, 12, 1)) 

ss3 <- subset(final_death_2015_2022, year == "2019") %>%
  select(month, drug_overdose, intentional_self_harm_suicide, assault_homicide) %>%
  gather(key = "variable", value = "value", -month)
p3 <- ggplot(ss3, aes(x = month, y = value)) + 
  geom_line(aes(color = variable), size = 1) +
  scale_color_manual(values = c("#1d95bb", "#0e9731", "#FC4E07")) +
  theme_minimal() +
  xlab("Month of Year, 2019") +
  ylab("Number of Deaths") +
  ggtitle("Number of Deaths by Select Cause, Year 2019") +
  scale_y_continuous(breaks=seq(0, 10000, 2000)) +
  scale_x_continuous(breaks=seq(1, 12, 1)) 

grid.arrange(p3, p2, p1, ncol = 1)
```

### **Save File as RDS for Group Work**
*Commenting out RDS creation so this is not overwritten*
RDS file `final_death_2015_2022.rds` located in `\nicoleluisi-MADA-portfolio\Mod_4_Exercise\finaldata`
```{r}
#saveRDS(final_death_2015_2022, file = (here("Mod_4_Exercise", "finaldata", "final_death_2015_2022.rds")))
```


###### Summary table
```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS(here("Mod_4_Exercise", "finaldata", "final_death_2015_2022.rds"))
knitr::kable(resulttable)
```
