---
title: "Module 8: Exploration"
author: "Nicole Luisi"
date: "`r Sys.Date()`"
output: html_document
---

This Quarto file was developed for the MADA Module 8 exercise.

### **Load and Process Data**

##### **Load required packages**

```{r}
#| warning: false
library("here")
library("tidyverse")
library("gt")
library("ggplot2")
library("reshape2")
library("grid")
library("gridExtra")
```

##### **Load data**

```{r}
# Note: Ignore extension on file, actually an RDS file 
mod8 <- readRDS(here("fluanalysis", "data", "mod8_clean.rds"))
```

##### **Explore data**

##### For each (important) variable, produce and print some numerical output (e.g. a table or some summary statistics numbers).
    
```{r}
# Quick look at data
names(mod8)
str(mod8)
glimpse(mod8)

mod8 %>% gt::gt() %>% 
  gt::tab_header(title = "Flu Data") %>%
  gt::tab_style(style = list(gt::cell_fill(color = "#7cb9e8"),
                             gt::cell_text(weight = "bold")),
                locations = gt::cells_body(columns = c(BodyTemp, Nausea)))

# Quick look at stats for key outcome vars
summary(mod8$BodyTemp)
table(mod8$Nausea, useNA="always")

# Separate yes/no vars and 4-level severity vars into subsets 
yesno_vars<-mod8 %>% select(SwollenLymphNodes, ChestCongestion, ChillsSweats, NasalCongestion, CoughYN, Sneeze, Fatigue, SubjectiveFever, Headache,         
                            WeaknessYN, CoughYN2, MyalgiaYN, RunnyNose, AbPain, ChestPain, Diarrhea, EyePn, Insomnia, ItchyEye, Nausea, EarPn, Hearing,
                            Pharyngitis, Breathless, ToothPn, Vision, Vomit, Wheeze)
sevlev_vars<- mod8 %>% select(Weakness, CoughIntensity, Myalgia)

# Look at counts for all yes/no vars
levs2 <- c("No", "Yes")
yesno_counts<-t(data.frame(sapply(yesno_vars, function(x) table(factor(x, levels=levs2, ordered=TRUE)))))
yesno_counts
 
# Look at counts for all 4-level severity vars
levs4 <- c("None", "Mild", "Moderate", "Severe")
sevlev_counts<-t(data.frame(sapply(sevlev_vars, function(x) table(factor(x, levels=levs4, ordered=TRUE)))))
sevlev_counts               
```

##### For each (important) continuous variable, create a histogram or density plot.
    
```{r}
ggplot(mod8, aes(mod8$BodyTemp)) +
  geom_histogram(aes(y = ..density..), fill='lightblue', col='blue', bins=50) +
  labs(title = "Histogram for Body Temperature",
       x = "Body Temperature",
       y = "Frequency") +
  stat_function(fun = dnorm, args = list(mean=mean(mod8$BodyTemp), sd=sd(mod8$BodyTemp)), col='red', lwd=1, lty='dashed') 
```

##### Create scatterplots or boxplots or similar plots for the variable you decided is your main outcome of interest and the most important (or all) independent variables/predictors (choose a few predictor variables).

```{r}
# Outcome 1: Body Temp
boxplot(mod8$BodyTemp,
main = "Mean Body Temperature",
xlab = "Body Temperature",
ylab = " ",
col = "lightblue",
border = "red",
horizontal = TRUE,
notch = TRUE
)

# Outcome 2: Nausea
ggplot(mod8, aes(x=reorder(Nausea, Nausea, function(x)-length(x)))) +
  geom_bar(fill='lightblue') +
  labs(x='Nausea', y='Frequency')

# Outcome 1 and 2 together
ggplot(mod8, aes(x=Nausea, y=BodyTemp)) +
  labs(title = "Box Plot of Body Temp with Nausea") +
  geom_boxplot(fill='lightblue')

# Look at some predictors with each outcome

# BodyTemp
#df.m <- melt(mod8, id.var = "Label")
p1<-ggplot(mod8, aes(x=SwollenLymphNodes, y=BodyTemp)) +
  geom_boxplot(fill='steelblue')
p2<-ggplot(mod8, aes(x=NasalCongestion, y=BodyTemp)) +
  geom_boxplot(fill='red')
p3<-ggplot(mod8, aes(x=SubjectiveFever, y=BodyTemp)) +
  geom_boxplot(fill='orange')
p4<-ggplot(mod8, aes(x=Diarrhea, y=BodyTemp)) +
  geom_boxplot(fill='green')
p5<-ggplot(mod8, aes(x=Vomit, y=BodyTemp)) +
  geom_boxplot(fill='purple')
p6<-ggplot(mod8, aes(x=Pharyngitis, y=BodyTemp)) +
  geom_boxplot(fill='yellow')
p7<-ggplot(mod8, aes(x=CoughYN, y=BodyTemp)) +
  geom_boxplot(fill='lightblue')
p8<-ggplot(mod8, aes(x=EarPn, y=BodyTemp)) +
  geom_boxplot(fill='pink')
p9<-ggplot(mod8, aes(x=AbPain, y=BodyTemp)) +
  geom_boxplot(fill='grey')
grid.arrange(p1,  p2,  p3,  p4,  p5,  p6,  p7, p8, p9,
  nrow = 3,
  top = "Body Temp with Select Predictors",
  bottom = textGrob(".",
    gp = gpar(fontface = 3, fontsize = 9),
    hjust = 1,
    x = 1
  )
)

# Nausea
n_Weakness <- table(mod8$Weakness,
                   mod8$Nausea)
barplot(t(n_Weakness),
        legend.text = TRUE,
        beside = TRUE,
        xlab="Weakness", args.legend=list(title="Nausea"), col=1:2)

n_CoughIntensity <- table(mod8$CoughIntensity,
                   mod8$Nausea)
barplot(t(n_CoughIntensity),
        legend.text = TRUE,
        beside = TRUE,
        xlab="Cough Intensity", args.legend=list(title="Nausea"), col=1:2)

n_SubjectiveFever <- table(mod8$SubjectiveFever,
                   mod8$Nausea)
barplot(t(n_SubjectiveFever),
        legend.text = TRUE,
        beside = TRUE,
        xlab="Subjective Fever", args.legend=list(title="Nausea"), col=1:2)
```

