---
title: "Module 8: Model Fitting"
author: "Nicole Luisi"
date: "`r Sys.Date()`"
output: html_document
---

This Quarto file was developed for the MADA Module 8 exercise.

### **Load and Process Data**

##### **Load required packages**

```{r}
#| warning: false
library(here)
library(tidyverse)
library(ggplot2)
library(tidymodels)
library(dplyr)
```

##### **Load data**

```{r}
# Note: Ignore extension on file, actually an RDS file 
mod8 <- readRDS(here("fluanalysis", "data", "mod8_clean.rds"))
```

##### **Fit a linear model to the continuous outcome (BodyTemp) using only the main predictor of interest (RunnyNose).**

```{r}
# Simple model with BodyTemp ~ RunnyNose
lm_mod <- linear_reg()
lm_mini <- 
  lm_mod %>% 
  fit(BodyTemp ~ RunnyNose, data = mod8)
tidy(lm_mini)
```

##### Fit another linear model to the continuous outcome using all (important) predictors of interest.

```{r}
lm_mod <- linear_reg()
lm_full <-  lm_mod %>%  set_engine("lm") %>% 
  fit(BodyTemp ~ . ,data = mod8)
tidy(lm_full)
```

##### Compare model results for the model with just the main predictor and all predictors.

```{r}
anova(lm_mini$fit, lm_full$fit)

glance(lm_mini) %>%
  dplyr::select(adj.r.squared, sigma, AIC, BIC, p.value)

glance(lm_full) %>%
  dplyr::select(adj.r.squared, sigma, AIC, BIC, p.value)
```

##### Fit a logistic model to the categorical outcome (Nausea) using only the main predictor of interest.

```{r}
lrg_mini <- logistic_reg() %>% 
  set_engine("glm") %>% 
  fit(Nausea ~ RunnyNose, data = mod8)
tidy(lrg_mini)
```

##### Fit another logistic model to the categorical outcome using all (important) predictors of interest.

```{r}
lrg_full <- logistic_reg() %>% 
  set_engine("glm") %>% 
  fit(Nausea ~ ., data = mod8)
tidy(lrg_full)
```

##### Compare model results for the categorical model with just the main predictor and all predictors.

```{r}
anova(lrg_mini$fit, lrg_full$fit)
```